Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-06 submission.tex
summary

\section{Summary and Future Work}
\label{sec:summary}
The improvements in parser performance with our optimized tag set indicate that
a more fine-grained PoS tag set may assist syntactic parsers in recognizing and
generalizing syntactic patterns, while potentially compromising the performance
of PoS taggers.

There are several aspects of this thesis that can be further explored in future
work, including extrinsic evaluation of the effects of PoS tag sets on other
downstream NLP applications besides parsing, such as sentiment analysis and
named entity recognition. These applications often require tagged data, but are
markedly different from syntactic parsing, hence the evaluation would involve
investigating an entirely different aspect of the effects of tag set
granularity.

>>>>

\section{Summary}
\label{sec:summary}
This paper has reported on a range of experiments with injecting more
fine-grained morphological distinctions into an existing PoS tag set, and then
empirically evaluating the effects both (intrinsically) in terms of tagging
accuracy and (extrinsically) in terms of parsing accuracy. Our experimental
results -- based on the annotations of the Norwegian Dependency Treebank using
the TnT PoS tagger and the Mate dependency parser -- show that the enriched tag
set leads to significantly improved parsing accuracy, even though PoS tagging
accuracy in isolation is reduced. The modified tag sets in our experiments are
defined as combinations of PoS tags and morphological features already
available in the gold annotations of the treebank. The best performing tag
configuration is in effect a PoS tag set optimized for dependency parsing of
Norwegian. However, by using another task for the extrinsic evaluation one can
of course optimize a tag set specifically for other downstream applications
instead, such as sentiment analysis, named entity recognition or any other task
making use of tagged data.
